### 2017-07-25 Writing/draft.tex
rewritten intro + methods

The last five years have seen a fantastic expansion of the use of discrete morphological data in macroevolutionary studies.
In fact, traditionally confined to ``classic'' cladistic studies \citep{felsenstein2004inferring}, these methods have recently been widely expended and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep{pyrondivergence2011,ronquista2012}.
In parallel, much development has also been done on pre-phylogenetic analysis \citep[e.g. data collection;][]{morphobank} as well as on post-phylogenetic analysis \citep[e.g. morphological disparity analysis;][]{Close2015,Claddis}.
Nonetheless, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146,bapst2017combined}; the coding of this data \citep{Brazeau2011}; the inference \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted} or the dating methods \citep{Arcila2015131,o2016tips}.

One aspect of the data, however, is often neglected: the correlation between morphological characters.
In fact, the assumption of character independence is at the base of phylogenetic inference and as been mentioned or implied throughout the history of phylogenetic methods \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
Nowadays, this assumption of character independence is even often accepted as a fact and rarely tested or questioned in empirical datasets.
However, specifically in the case of discrete morphological data, this assumption of independence is probably violated more often than not due to the sheer nature of the data. In practice, dependence between characters can be induced in at least three different ways:

\begin{itemize}
    \item \textbf{Biological induce dependence between characters:} this is the result of a link between two characters through development or biological function (e.g. upper and lower teeth).
    These links are readily understood through developmental biology or understanding of shared function between features \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
    One trivial example would be the link between lower and upper molar characters in mammals: one character describing a feature of a lower molar will be expected to be correlated to another character describing a feature of the associated upper molar.

    \item \textbf{Evolutionary induced dependence between characters:} this is the result of sets of character co-evolving along branches in the phylogeny.
    The character's correlation is thus due to shared evolutionary history of these characters.
    Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
    For example, in gnatosthomes, the apparition of the jaw in the gnatosthomes clade correlates with a whole sweep of characters present in modern gnatosthomes (e.g. beaks, tribosphenic molars, jaw protrusion, etc.). %TG: this links to inapplicable data!

    \item \textbf{Coding induced dependence between characters:} this is the results of researchers methodology for defining or/and coding discrete morphological characters \citep[see][and reference therein]{Brazeau2011}.
    These dependences could be due to methodological errors. 
    For example one could use a character describing a particular feature on a left molar as well as another character describing the same feature on a right molar.
    However, these dependences could also be due to the nature of the available data, especially in palaeontology.
    For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
\end{itemize}

\noindent Of course, the three sources of dependences can also be correlated to each other: characters describing the left and right lower/upper molars will have induce correlation due to the modularity of the molars, their shared history and the duplicated coding!

@@

\noindent \textbf{Software induced correlation between characters:} this is the result how software actually interprets the differences between characters.
I.e. the vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treated as different/similar tokens) leading to dependence between characters.
For example, in a matrix containing four cetartiodactyles - say a pig (e.g. \textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four characters - say (\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragal with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley - resulting in the following matrix:

@@

In the example in Table \ref{Tab:example_matrix}, the characters C1 and C2 are the most likely to be truly independent; characters C3 and C4 have a coding induced dependency; characters C1 and C3/C4 have an evolutionary induce dependency and again, characters C2 and C3/C4 are likely to be independent.
Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the different character states tokens will be used in order to infer a phylogenetic tree.

To our knowledge, this effect of software induced character correlation

It is thus still unclear what will be the effect of characters correlation on phylogenetic inference: how does these correlation really affect topology?
Additionally, because of the nature of discrete morphological characters and organisms morphology, correlation is expected to rise through both combinatorics properties of the characters or simply through evolutionary processes.
What is the effect of these two sources of correlations and can we (should we) avoid them?

In this study, we formally assess the effect of discrete character's correlation using simulated and empirical data.
We propose a new metric to measure the correlation (or difference) between two discrete characters and a protocol to modify discrete data matrices to increase/decrease the correlation between characters within/
We found that blabalbal

@@

We applied the following simulation protocol to assess the effect of characters correlation (Fig.\ref{Fig:outline}):
\begin{enumerate}
    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by duplicating characters in order to maximise and minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters. We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximised or minimised character difference to create null expectancy matrices (hereafter called the ``null'' matrices).
    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``null'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``null'' trees).
    \item \textbf{Comparing topologies}: finally, we compared to topologies in two ways: (1) first we compared the ``normal'' to the ``null'' trees to assess our simulation protocol; (2) second, we compared the ``maximised'' and ``minimised'' trees to the ``null'' and ``normal'' trees respectively to measure the signal to noise ratio and thus measure the effect of character correlation on topology.
\end{enumerate}
Each steps are described below in more details, as well as our proposed definition of the Character Difference metric.

@@

\subsection{Measuring correlation between characters}
We define characters as being correlated if they give the same phylogenetic information.
In order to measure this similitude in phylogenetic information, we propose a new distance metric to measure the difference between two characters:

\subsubsection{Characters difference (CD)}
\begin{equation}
    CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
\end{equation}

\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
$CD$ is a continuous distance metric valid in a mathematical sense (see Supplementary).
Since we are considering differences as being only Fitch-like (i.e. non-weighted and non-ordinate), we ignore the actual distance between two different character states tokens \citep[i.e. a Gower distance;][]{GowerDist}.
Two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
In order to facilitate and standardise characters states comparison, we standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
In other words, we replaced all the occurrences of the first token to be $0$, the second to be $1$, etc.
This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{0,0,1,2,2,1\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).
When the character difference is equal to $0$, it means that characters convey the same phylogenetic signal.
When the character difference is equal to $1$ it means it conveys the most different signal.
For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.

\subsection{Simulating discrete morphological matrices}
To simulate the matrices we applied a protocol very similar to \citep{Guillerme2016146}.
First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform ($0$,$1$) distribution maintaining $\lambda$ $>$ $\mu$ using the diversitree \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
For each tree, we arbitrarily set the outgroup to be the first taxa (alphabetically) thus effectively rooting the trees on this taxa.
These trees are hereafter called the ``true'' trees.
We then simulated discrete morphological characters using the either of the two following models:
\begin{itemize}
    \item The HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ an scale to sum to one) and using a transition/transvertion rate of two \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][; see below]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
    \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146}.
    This model assumes a equal transition rate between character states that might seem overly simplifying and excludes other observed transition patterns \citep[e.g. Dollo characters;][]{Dollo,wright2015came}.
    Recently however, \cite{Wright01072016} have shown that an equal rate transition is still strictly majority in empirical data.
\end{itemize}

\noindent For each character, both models (HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
This low evolutionary rate values allowed to reduce the number of homoplasic character changes and thus reinforce the phylogenetic information in the matrices (see supplementary materials @@@).
We re-simulated every invariant characters to obtain a matrix with no invariant characters.
To ensure that our simulation where reflecting realistic observed parameters, we only selected matrices with Consistency Indices superior to $0.26$ \citep{OReilly20160081}.

For each trees with 25, 75 or 150 taxa we generated matrices with 100, 350 and 1000 characters following \cite{OReilly20160081}.
The matrices where generated using the \texttt{dispRity R} package \citep[][; \url{https://github.com/TGuillerme/dispRity}]{thomas_guillerme_2016_55646}.
To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 50 times resulting in 450 morphological matrices (hereafter called the ``normal'' matrices).

\subsection{Modifying the matrices}

We calculated the pairwise character differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}. %TG: TO IMPLEMENT! 
We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e. with an average character difference $<$$0.25$) and replaced them randomly by any of the remaining characters.
This operation increased the overall pairwise character difference in the matrix thus making the characters more dissimilar.
Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (i.e. with an average character difference $<$$0.75$) and randomly replaced them with the remaining ones.
Finally, because this operation effectively changes the weight of characters (i.e. giving the characters $<$$0.25$ or $>$$0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a null expectation modification (i.e. randomly weighting characters).
This step resulted in a total of 1800 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``null'' matrices).
The algorithms for the three modifications are available on GitHub@@@. %TG: "@@@" is my "to replace" sign.

\subsection{Inferring topologies}

We inferred the topologies in both Bayesian and Maximum Parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
The Maximum Parsimony inference was run over @@@ CPU hours using a heuristic search with random sequence addition replicate 10 times (\texttt{hsearch addseq=random nreps=10}).
The trees where also bootstrapped 100 times using the same search parameters mentioned above.

The Bayesian inference was run over @@@ CPU hours

% lset nst=1 rates=gamma Ngammacat=4;
% prset ratepr=variable Shapepr=Exponential(0.5);

% mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES file=150t_100c_0304_maxi;
% sump Filename=150t_100c_0304_maxi Relburnin=YES Burninfrac=0.25;
% sumt Filename=150t_100c_0304_maxi Relburnin=YES Burninfrac=0.25;

 % "hsearch addseq=random nreps=500 rseed=01234 rearrlimit=5000000;"

Due to cluster hardware requirements an to save some time, when chains didn't converged and the runs exceeded 10GB, we aborted the MCMC and computed the consensus tree from the uncoverged chains.
In practice, these chains got stuck at an ASDS around (but not below) 0.01.

\subsection{Comparing topologies}

We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``null'' matrices and the tree inferred from the ``normal'' matrix.

We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the different inferred trees.
We ran two types of comparisons, (1) the ability to recover the correct tree (``normal'' trees \textit{vs} ``null'' and ``true'' trees and (2) the effect of character differences by measuring the noise to signal ratio between the ``maximised'' and ``minimised'' trees and the ``null'' and ``normal'' trees.

@@

% \subsection{Empirical morphological matrices}

% We applied the same procedure described above (without the morphological matrices simulation) to 100 empirical matrices downloaded from TreeBASE (\url{http://treebase.org/}).
% The 100 matrices where selected to be the most representative of empirical work with discrete morphological data: with randomly selected matrices with more than 100 each, published between 1985 and 2013 and covering 19 taxonomic classes within 6 phyla (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
% The full list of matrices is available in the supplementary materials @@@.

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/RF_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
\caption{Effect of Character Difference on recovering the ``best'' topology. The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively. The x axis represents the different Character Difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}
\label{Fig:RF_results_best}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
\caption{Effect of Character Difference on recovering the ``best'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_best}
\end{figure}

@@

@@

\section{Acknowledgements}
Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232).
We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD proof.

>>>>

The last six years have seen a fantastic expansion of the use of discrete morphological data in macroevolutionary studies.
Traditionally confined to ``classic'' cladistic studies, these methods have recently been widely expanded and improved whether based on the models and methods for inferring phylogenies \citep[e.g.][]{heath2014fossilized,Wright01072016} or to combine them with molecular data \citep[e.g.][]{pyrondivergence2011,ronquista2012}.
In parallel, much development has also been done on pre-phylogenetic analysis \citep[e.g. data collection;][]{morphobank} as well as on post-phylogenetic analysis \citep[e.g. morphological disparity analysis;][]{Close2015,Claddis}.
Nonetheless, using discrete morphological data in macroevolutionary studies comes with several caveats based to the nature of data used \citep{Guillerme2016146,bapst2017combined}; the coding of this data \citep{Brazeau2011,simoes2017giant}; the inference method \citep{spencerefficacy2013,wrightbayesian2014,OReilly20160081,puttick2017uncertain,goloboff2017weighted} or the dating methods \citep{Arcila2015131,o2016tips}.

One aspect of the data, however, is often neglected: the correlation between morphological characters.
In fact, the assumption of character independence is at the base of phylogenetic inference and as been mentioned or implied throughout the history of phylogenetic methods \citep[e.g.][]{joysey1982problems,felsenstein1985phylogenies,lewisa2001,felsenstein2004inferring}.
Nowadays, this assumption of character independence is nearly always accepted as a fact and rarely tested or questioned in empirical datasets.
However - especially for discrete morphological data - this assumption of independence is probably violated more often than not due to the sheer nature of phylogenetic data.
In practice, dependence between characters can be induced in at least three different ways:

\begin{itemize}
    \item \textbf{Biological induce dependence between characters:} this is the result of a link between two characters through development and/or biological function.
    These links are readily understood through developmental biology or understanding of shared function between features \citep{goswami2006morphological,goswami2010,goswami2014macroevolutionary}.
    One trivial example would be the link between lower and upper molar characters in mammals: one character describing a feature of a lower molar will be expected to be correlated to another character describing another feature of the associated upper molar.

    \item \textbf{Evolutionary induced dependence between characters:} this is the result of sets of characters co-evolving along branches in the phylogeny.
    The character's correlation is thus due to shared evolutionary history of these characters.
    Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
    For example, the apparition of the jaw as a character in gnatosthomes correlates with a whole sweep of characters present in modern gnatosthomes (e.g. beaks, tribosphenic molars, jaw protrusion, etc.). %TG: this links to inapplicable data!

    \item \textbf{Coding induced dependence between characters:} this is the results of researchers methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
    These dependences could be due to methodological errors. 
    For example one could define a character describing a particular feature on a left molar as well as another character describing the same feature on a right molar.
    It is worth noting that these dependences could also be due to the nature of the available data, especially in palaeontology.
    For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
\end{itemize}

\noindent Of course, the three sources of dependences can also be correlated to each other: characters describing the left and right lower/upper molars will have induced dependence due to the modularity of the molars, their shared history and the duplicated coding!

@@

\noindent \textbf{Software induced correlation between characters:} this is the result how software actually interprets the differences between characters.
I.e. the vast majority of phylogenetic software ignores both the character's definition and the different states signification (simply treating them as different/similar tokens) leading to dependence between characters.
For example, in a matrix containing four cetartiodactyles - say a pig (e.g. \textit{Sus}), a deer (\textit{Cervus}), a hippo (\textit{Hippopotamus}) and a whale (\textit{Balaenoptera}) - and four characters - say (\textbf{C1}: presence (1) or absence (0) of an astragal; \textbf{C2}: presence (0) or absence (1) of baleens; \textbf{C3}: presence (0) or absence (1) of a left astragal with a double pulley; \textbf{C4}: presence (0) or absence (1) of a right astragal with a double pulley - resulting in the following matrix:

@@

In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the different character states tokens will be used in order to infer a phylogenetic tree.
It is thus still unclear what will be the effect of characters correlation on phylogenetic inference: how does these correlation really affect topology?

If the three induced sources of correlation are fairly well studied, their actual interpretation in phylogenetic inference, has, to our knowledge, never been studied through a simulation protocol.
In this study, we formally assess the effect of discrete character's correlation using simulated and empirical data.
We propose a new metric to measure the difference between characters (as a proxy for these three sources of correlation) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences/similarities between characters.
We found that blabalbal

@@

We applied the following simulation protocol to assess the effect of characters correlation (Fig.\ref{Fig:outline}):
\begin{enumerate}
    \item \textbf{Simulating matrices}: we simulated discrete morphological matrices with 25, 75 and 150 taxa for 100, 350 and 1000 characters, hereafter called the ``normal'' matrices.
    \item \textbf{Modifying matrices}: we modified the ``normal'' matrices by duplicating characters in order to maximise or minimise characters differences (hereafter called respectively ``maximised'' and ``minimised'' matrices) by removing respectively the least or most different characters and replacing them randomly by the remaining characters.
    We also randomly duplicated characters from the ``normal'' matrices without biasing towards maximising or minimising character difference to create randomised matrices (hereafter called the ``randomised'' matrices - equivalent to a null expectancy).
    \item \textbf{Inferring topologies}: we inferred the topologies from the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices using both Maximum Parsimony and Bayesian inference (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' trees).
    \item \textbf{Comparing topologies}: finally, we compared to topologies in two ways: (1) first we compared the ``normal'' to the ``randomised'' trees to assess our simulation protocol; (2) second, we compared the ``maximised'' and ``minimised'' trees to the ``randomised'' and ``normal'' trees respectively to measure the effect of character correlation on topology.
\end{enumerate}
Each steps are described below in more details, as well as our proposed definition of the Character Difference metric.

@@

\subsection{Measuring differences between characters}
We define characters as being correlated if they give the same phylogenetic information.
In order to measure this similitude in phylogenetic information, we propose a new distance metric to measure the difference between two characters:

\subsubsection{Characters Difference (CD)}
\begin{equation}
    CD_{(x,y)} = 1 - \left(\frac{\abs*{\frac{\sum_{i}^{n}\abs{x_{i} - y_{i}}}{n}-\frac{1}{2}}}{\frac{1}{2}}\right)
\end{equation}

\noindent Where $n$ is the number of taxa with comparable data and $x_i$ and $y_i$ are each characters states for the characters $x$ and $y$ and the taxa $i$.
$CD$ is a continuous distance metric valid in a mathematical sense (see Supplementary Material 1).
Since we are considering differences as being only Fitch-like (i.e. non-weighted and non-ordinate), we calculated the difference between character states in a qualitative way.
Two same character states tokens have a difference of $0$ and two different ones have a difference of $1$ (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
In order to facilitate and standardise characters states comparison, we standardised each characters by arbitrarily modifying their character states tokens by order of appearance.
In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
This way, a character \texttt{A = \{2,2,3,0,0,3\}} would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \citep[following the \textit{xyz} notation in][p.13]{felsenstein2004inferring}.
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct groups).

When the character difference is equal to $0$, it means that characters convey the same phylogenetic signal.
When the character difference is equal to $1$ it means it conveys the most different signal.
For example with three characters \texttt{A = \{0,1,1,1\}}, \texttt{B = \{1,0,0,0\}} and \texttt{C = \{0,1,2,3\}}, $CD_{(A,B)} = 0$ and $CD_{(A,C)} = 1$.

\subsection{Simulating discrete morphological matrices}
To simulate the matrices we applied a protocol very similar to \citep{Guillerme2016146}.
First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform ($0$,$1$) distribution maintaining $\lambda$ $>$ $\mu$ using the diversitree \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
For each tree, we arbitrarily set the outgroup to be the first taxa (alphabetically) thus effectively rooting the trees on this taxa.
These trees are hereafter called the ``true'' trees.
We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
\begin{itemize}
    \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a uniform distribution $(0,1)$ and scaled to sum to $1$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use a M$k$ model;][; see below]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
    \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146}.
    This model assumes a equal transition rate between character states that might seem overly simplifying and excludes other observed transition patterns \citep[e.g. Dollo characters;][]{Dollo,wright2015came}.
    Recently however, \cite{Wright01072016} have shown that an equal rate transition is still the most present in empirical data.
\end{itemize}

\noindent For each character, both models (morphological HKY-binary or M$k$) where chosen randomly and ran with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
This low evolutionary rate values allowed to reduce the number of homoplasic character changes and thus reinforce the phylogenetic information in the matrices.
We re-simulated every invariant characters to obtain a matrix with no invariant characters.
To ensure that our simulation where reflecting realistic observed parameters, we only selected matrices with Consistency Indices superior to $0.26$ \citep{OReilly20160081}.

It can be noted that this Consistency Index threshold might favor Maximum Parsimony inference (since we measured this index using quick parsimony trees) yet the evolutionary models of each character is slightly in favor of Bayesian inference (since, on average, half the characters evolved following a M$k$ model).

For each trees with 25, 75 or 150 taxa we generated matrices with 100, 350 and 1000 characters following \cite{OReilly20160081}.
The matrices where generated using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}.
To estimate the variance of our simulations and assess the effect of our random parameters, we repeated this step 40 times resulting in 360 ``normal'' morphological matrices.

\subsection{Modifying the matrices}
We calculated the pairwise Character Differences for each generated matrices using the \texttt{dispRity R} package \citep{thomas_guillerme_2016_55646}. %TG: TO IMPLEMENT! 
We then modified the matrices to either maximise or minimise the pairwise character differences for each matrices using three different algorithms.
For maximising the pairwise differences between characters, we selected the characters that where the most similar to all the others (i.e. with an average character difference $<$$0.25$) and replaced them randomly by any of the remaining characters.
This operation increased the overall pairwise character difference in the matrix thus making the characters more dissimilar.
Conversely, for minimising the pairwise character differences, we selected the most dissimilar characters (i.e. with an average character difference $<$$0.75$) and randomly replaced them with the remaining ones.
Finally, because this operation effectively changes the weight of characters (i.e. giving the characters $<$$0.25$ or $>$$0.75$ a weight of $0$ and giving the randomly selected remaining characters a weight of +$1$), we randomly replaced the average number of characters replaced in the character maximisation and minimisation by any other characters as a randomised expectation modification (i.e. randomly weighting characters).
This step resulted in a total of 1440 matrices (hereafter called the ``normal'', ``maximised'', ``minimised'' and ``randomised'' matrices - see Fig. \ref{Fig:modif_matrix} for an illustration).
The algorithms for the three modifications are available on GitHub@@@. %(Modifying_matrices.html) %TG: "@@@" is my "to replace" sign.

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Modif_matrix.pdf}
\caption{Example illustration of the protocol for modifying matrices. The matrices represent the pairwise character differences for 100 characters. Blue colours correspond to low character differences and orange colours correspond to high character differences. \textbf{A} - a random Birth-Death tree is simulated and used for generating the normal matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised character difference (\textbf{C}), minimised it (\textbf{D}) or randomise it (\textbf{E}).}
\label{Fig:modif_matrix}
\end{figure}

\subsection{Inferring topologies}
We inferred the topologies in both Bayesian and Maximum Parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
For both methods, we set the arbitrarily chosen outgroup as an outgroup in the models.
The Maximum Parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).

The Bayesian inference was run using an M\textit{k} model with ascertainment bias and four discrete gamma rate categories (M\textit{kv} $4\Gamma$ - \texttt{lset nst=1 rates=gamma Ngammacat=4}) with an variable rate prior an exponential (0.5) shape (\texttt{prset ratepr=variable Shapepr=Exponential(0.5)}).
We ran two runs of 6 chains each (2 hot, 4 cold) for a maximum of $1\times10^9$ generations with a sampling every 200 generations.
We automatically stopped the MCMC when the average standard deviation of split frequencies (ASDSF) between both runs fell below 0.01 (with a diagnosis every $1\times10^4$ generations - \texttt{mcmc nruns=2 Nchains=6 ngen=1000000000 samplefreq=200 printfreq=2000 diagnfreq=10000 Stoprule=YES stopval=0.01 mcmcdiagn=YES}).
Due to cluster hardware requirements an to save some time, when chains didn't converged and the runs exceeded 5GB each, we aborted the MCMC and computed the consensus tree from the uncoverged chains.
In practice, these few MCMC got stuck at an ASDSF around (but not below) 0.01.

A strict majority rule tree was then calculated for both Bayesian an Maximum Parsimony trees.
For the Bayesian consensus trees, the 25\% first trees of the posterior tree distribution were excluded as a burnin.
The 2880 tree inferences took around 1 CPU century on the Imperial College High Performance Computing Service \citep[2-3GHz clock rate;][]{HPC}.

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Fould distance \citep{RF1981} and the Triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We ran two types of comparisons, (1) the effect of Character Difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}); and (2) the signal to noise ratio induced by the simulation protocol by comparing the ``randomised'' trees to the ``normal'', ``maximised'' and ``minimised'' trees (Figs \ref{Fig:RF_results_rand} and \ref{Fig:Tr_results_rand}).

@@

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
\caption{Effect of Character Difference on recovering the ``best'' topology. The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively. The x axis represents the different Character Difference scenarios and tree inference method with the Maximised Character Difference in Bayesian (red) and under Maximum Parsimony (orange), the Minimised Character Difference in Bayesian (dark green) and under Maximum Parsimony (light green) and the Randomised Character Difference in Bayesian (dark blue) and under Maximum Parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}
\label{Fig:RF_results_best}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf} %TG: These figures will be cleaned up once I have the final results.
\caption{Effect of Character Difference on recovering the ``best'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_best}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/RF_results_null.pdf} %TG: These figures will be cleaned up once I have the final results.
\caption{Effect of Character Difference on recovering the ``random'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Ronbinson-Fould distance.}
\label{Fig:RF_results_rand}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Tr_results_null.pdf} %TG: These figures will be cleaned up once I have the final results.
\caption{Effect of Character Difference on recovering the ``random'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_rand}
\end{figure}

@@

This goes with former results (see Parins-Fukuchi in review.)

@@

\section{Acknowledgements}
Calculations where done using the Imperial College London Cluster Services (doi: 10.14469/hpc/2232).
We thank Alberto Pascual Garcia for input in the design of the simulations protocol and Guillermo Herraiz Yebes for helping with the CD metric proof.
